%% acknowledgments.tex

% From mitthesis package
% Version: 1.02, 2024/06/19
% Documentation: https://ctan.org/pkg/mitthesis

\chapter*{Acknowledgments}
\pdfbookmark[0]{Acknowledgments}{acknowledgments}


The completion of this thesis is a minor miracle only brought about through the help of many others. Put simply, I would not have been able to accomplish this work without help from many others, not just this last semester but very much over the last four years of my undergraduate education at MIT. While no list like this can be truly exhaustive, I want to acknowledge as many of these people as I can.

\textbf{Family: }
My family deserves the most appreciation in the world. Throughout the rollercoaster of a time that is a MIT education, they have loved and supported me unconditionally, pushed me to do my best, and have always lent an ear when I was struggling with something, whether it be personal or technical. I would also like to acknowledge the late Marc Perkel in particular, who recognized my love of engineering early and stoked that passion from an early age.

\textbf{Professors: }
First, I would like to acknowledge my thesis supervisor, Dr. Barbara Hughey. Dr. Hughey is one of the best educators I know: her class is both hands-on and educational, the perfect combination for learning. Much of the measurement methods I develop in this paper were first learned in 2.671.

I would also like to acknowledge professor Martin Culpepper, teacher of 2.72, for providing a fantastic senior capstone class in mechanical engineering that pushed me to be the best mechnaical engineer I can be. In short, this class truly changed the way I think and that is the ultimate compliment for any class.

\textbf{Experts: }
I would also like to acknowledge the work of many experts who made this work possible. First, to John Hopkins for his FACT method for flexture design, which was instrumental for 2.72 and for this work. Next, to Larry Howell for the excellent textbook on flexture design.
I would also like to acknowledge Prof. Mengjia Yan at MIT, and especially graduate students Vincent Ulitzch and Joseph Ravichandran for their support on the project. They are true experts in hardware security and introduced me to a wealth of new ideas that expanded my knowledge in the space.
Lastly, to my mentors personally, professionally and beyond - Len Shustek, Ephraim Nowak at Rain, Brad Haas at Tesla, Andrew Weiner and Zach Jauvtis at SpaceX especially, thank you for believing in me and giving me a range of projects to expand my skills in innumerable ways.

\textbf{Makerspaces: }
I would also like to thank the people of many machine shops and makerspaces at MIT that have supported a huge variety of projects over the last 3 years. This includes LMP (Laboratory for Manufacturing and Productivity) - Wade Warman, Daniel Gilbert and N52 (Patrick McAtamney) especially.

In particular, MAD (The Morningside Academy for Design) is home to two of the most helpful makers at MIT - Bill and Chris. Both went above and beyond many times to support me, and many others with personal projects at the expense of their own time and resources.

\textbf{Friends: }
During the last four years, I have also benefited from some incredible friendships that have pushed me to be a better person. On MIT Motorsports, I want to give thanks to Henry Smith, Megan Gupta-She, Hezekiah Pendley, Sam Huelsnitz, Keiji Imai, Ander Jurs, David Lopez for working on some of the hardest problems of my life. I certainly hope we work together again someday. To my 2.009 and 2.72 team members, Via, Evan, Jono, David and Jorge, I appreciate the time we have shared working on some seriously ambitious projects.

I also owe thanks to some specific mentors and friends. To Levi Gershon, my peer and dear friend, who led me onto the MIT Motorsports team and SpaceX and is an inspiration both technically and professionally. 
I would like to thank Robbie Khazan: we have climbed many mountains and explored hidden valleys together, and taught me to listen closely while enjoying life to the fullest.
	
Last, but certainly not least, to the MechE office staff Christina Spinelli and Emma Dunn, thank you for advocating for this project during all stages of its conception and execution. And thank you most of all for the great MechE merchandise.

\break

\textit{To all hobbyists and fellow learners of the world, may we one day become graceful experts.}
- Joseph Ravichandran, B.S. Thesis, UIUC 2021

\textit{From this day forward, Mission Control will be described by two words: Tough and Competent.} - Gene Kranz, Apollo Mission Control, 1965